---
title: |
  Supplementary Material:
  Comparing antigenic distance metrics for influenza
format:
  docx:
    toc: true
    number-sections: true
bibliography: project-refs.bib
csl: plos-computational-biology.csl
execute:
  echo: false
  message: false
  warning: false
lang: en-US
editor_options:
  chunk_output_type: console
---

```{r}
#| label: setup
#| include: false

library(flextable)

options(knitr.graphics.rel_path = FALSE)

fit_flextable_to_page <- function(tbl, page_width = 6.5) {
	ft_out <- flextable::autofit(tbl)
	ft_width <- (dim(ft_out)$widths * page_width) /
		flextable::flextable_dim(ft_out)$widths
	flextable::width(ft_out, width = ft_width)
}
```

# Reproducibility Instructions

The analysis is organized as a `targets` pipeline. From the repository root,
restore the project environment with `renv::restore()` if needed, then run the
pipeline with:

```r
targets::tar_make()
```

The pipeline defaults to `FLU_TARGETS_MODE=test`, which uses reduced stochastic
settings for development checks. Publication-scale settings are selected by
setting `FLU_TARGETS_MODE=full` before running the pipeline. The Mantel
correlation table in the main manuscript is generated from
`cophenetic_mantel_summary`, written to
`results/derived/cophenetic-mantel-summary.rds`, and rendered as
`results/Tables/distance-mantel-table.rds`.

# Distance-Matrix Comparison Details

The primary manuscript table reports Mantel correlations between ML-tree
cophenetic distance and each candidate distance metric. These tests use only
unique off-diagonal strain pairs. Duplicate symmetric distances and diagonal
self-comparisons are excluded before estimating the observed correlation,
Bayesian bootstrap interval, and matrix-permutation p-value.

For the overall rows, permutation is stratified by subtype: H1N1 and H3N2 labels
are permuted within subtype, then the unique-pair vectors are concatenated for
the overall Mantel statistic. Bayesian bootstrap intervals are resampled over
strain units and are separate from the permutation p-values. We do not report
Bayesian-bootstrap tail probabilities as p-values because the permutation
p-values are the pre-specified apparent-sample test for matrix correspondence.

```{r}
#| label: tbl-mantel-settings
#| tbl-cap: |
#|   Mantel correlation computation settings. Seeds are row-specific so that
#|   each comparison and reporting scope is reproducible.

here::here("results", "derived", "cophenetic-mantel-summary.rds") |>
	readr::read_rds() |>
	dplyr::select(
		"Comparison",
		"Scope",
		"Permutations",
		"Bootstrap draws",
		"CI method",
		"Permutation seed",
		"Bootstrap seed"
	) |>
	flextable::flextable() |>
	fit_flextable_to_page()
```

```{r}
#| label: tbl-mantel-main-copy
#| tbl-cap: |
#|   Main Mantel correlation table reproduced for reference.

here::here("results", "Tables", "distance-mantel-table.rds") |>
	readr::read_rds() |>
	fit_flextable_to_page()
```

## Descriptive Pearson Correlations

The Pearson correlations in @tbl-pearson-correlations are descriptive summaries
of the same unique off-diagonal pair representation used for the Mantel
analysis. They are included to help interpret the approximate linear agreement
between distance scales, but they are not used as independent-observation
inference.

```{r}
#| label: tbl-pearson-correlations
#| tbl-cap: |
#|   Descriptive Pearson correlations between each distance metric and
#|   ML-tree cophenetic distance, overall and by subtype. Intervals are
#|   strain-level bootstrap intervals and should be interpreted descriptively.

here::here("results", "Tables", "descriptive-pearson-distance-correlation-table.rds") |>
	readr::read_rds() |>
	fit_flextable_to_page()
```

# References

::: {#refs}
:::
