---
title: |
  Supplementary Material:
  Comparing antigenic distance metrics for influenza
format:
  docx:
    toc: true
    number-sections: true
bibliography: project-refs.bib
csl: plos-computational-biology.csl
execute:
  echo: false
  message: false
  warning: false
lang: en-US
editor_options:
  chunk_output_type: console
---

```{r}
#| label: setup
#| include: false

library(flextable)

options(knitr.graphics.rel_path = FALSE)

fit_flextable_to_page <- function(tbl, page_width = 6.5) {
	ft_out <- flextable::autofit(tbl)
	ft_width <- (dim(ft_out)$widths * page_width) /
		flextable::flextable_dim(ft_out)$widths
	flextable::width(ft_out, width = ft_width)
}
```

# Reproducibility Instructions

The analysis is organized as a `targets` pipeline. From the repository root,
restore the project environment with `renv::restore()` if needed, then run the
pipeline with:

```r
targets::tar_make()
```

The pipeline defaults to `FLU_TARGETS_MODE=test`, which uses reduced stochastic
settings for development checks. Publication-scale settings are selected by
setting `FLU_TARGETS_MODE=full` before running the pipeline. The Mantel
correlation table in the main manuscript is generated from
`cophenetic_mantel_summary`, written to
`results/derived/cophenetic-mantel-summary.rds`, and rendered as
`results/Tables/distance-mantel-table.rds`.

# Distance-Matrix Comparison Details

The primary manuscript table reports Mantel correlations between ML-tree
cophenetic distance and each candidate distance metric. These tests use only
unique off-diagonal strain pairs. Duplicate symmetric distances and diagonal
self-comparisons are excluded before estimating the observed correlation,
Bayesian bootstrap interval, and matrix-permutation p-value.

For the overall rows, permutation is stratified by subtype: H1N1 and H3N2 labels
are permuted within subtype, then the unique-pair vectors are concatenated for
the overall Mantel statistic. Bayesian bootstrap intervals are resampled over
strain units and are separate from the permutation p-values. We do not report
Bayesian-bootstrap tail probabilities as p-values because the permutation
p-values are the pre-specified apparent-sample test for matrix correspondence.

```{r}
#| label: tbl-mantel-settings
#| tbl-cap: |
#|   Mantel correlation computation settings. Seeds are row-specific so that
#|   each comparison and reporting scope is reproducible.

here::here("results", "derived", "cophenetic-mantel-summary.rds") |>
	readr::read_rds() |>
	dplyr::select(
		"Comparison",
		"Scope",
		"Permutations",
		"Bootstrap draws",
		"CI method",
		"Permutation seed",
		"Bootstrap seed"
	) |>
	flextable::flextable() |>
	fit_flextable_to_page()
```

```{r}
#| label: tbl-mantel-main-copy
#| tbl-cap: |
#|   Main Mantel correlation table reproduced for reference.

here::here("results", "Tables", "distance-mantel-table.rds") |>
	readr::read_rds() |>
	fit_flextable_to_page()
```

## Subtype Contrast Analysis

For each distance metric, the primary subtype contrast is defined as the H3N2
Mantel correlation minus the H1N1 Mantel correlation. The Mantel correlations
use the same unique off-diagonal strain-pair representation as the main
distance-matrix table. Uncertainty is estimated with a Bayesian bootstrap over
strain units [@rubin1981]. Within each bootstrap draw, strain weights are
normalized separately within H1N1 and H3N2, pair weights are calculated as the
product of the two endpoint strain weights, and the weighted H3N2-minus-H1N1
contrast is recomputed.

The primary contrast answers a strain-sampling question: how uncertain is the
observed subtype difference if the strains are treated as sampled units from a
broader strain universe? We also report two sensitivity analyses that answer
different questions. The Fisher-z approximation [@fisher1921] treats the
unique strain pairs as independent observations, so it is expected to be
overconfident because each strain contributes to many pairs. The independent
Mantel permutation contrast permutes matrix labels within each subtype and
compares the observed contrast against a joint no-association null; it is useful
as a matrix-aware sensitivity check, but it is not a direct test of equality
between subtype-specific Mantel correlations. Differences among these
summaries can therefore be driven by pairwise non-independence, different
effective sample sizes across subtypes, and the fact that the permutation
sensitivity asks a different null-hypothesis question than the bootstrap
interval.

In practice, the Fisher-z sensitivity should usually look more decisive than
the primary bootstrap interval because it uses the 153 H1N1 and 210 H3N2
pairwise distances as if they were independent observations. The permutation
sensitivity may also differ from the bootstrap interval: for example, it can
flag large temporal or cartographic contrasts against a no-association null
while not flagging smaller contrasts between two already-high subtype-specific
matrix associations.

```{r}
#| label: tbl-subtype-contrast-settings
#| tbl-cap: |
#|   Subtype contrast bootstrap settings. Seeds are row-specific so that each
#|   contrast is reproducible.

here::here("results", "derived", "cophenetic-subtype-contrast-summary.rds") |>
	readr::read_rds() |>
	dplyr::select(
		"Comparison",
		"Bootstrap draws",
		"CI method",
		"Bootstrap seed"
	) |>
	flextable::flextable() |>
	fit_flextable_to_page()
```

```{r}
#| label: tbl-subtype-contrast-main-copy
#| tbl-cap: |
#|   Main subtype contrast table reproduced for reference.

here::here("results", "Tables", "subtype-mantel-contrast-table.rds") |>
	readr::read_rds() |>
	fit_flextable_to_page()
```

```{r}
#| label: tbl-subtype-contrast-sensitivity
#| tbl-cap: |
#|   Sensitivity summaries for H3N2-minus-H1N1 subtype contrasts. These
#|   alternatives are included to show how the contrast behaves under simpler
#|   Fisher-z and matrix-permutation reference calculations; neither replaces
#|   the primary strain-bootstrap contrast interval.

here::here("results", "Tables", "subtype-mantel-contrast-sensitivity-table.rds") |>
	readr::read_rds() |>
	fit_flextable_to_page()
```

## Descriptive Pearson Correlations

The Pearson correlations in @tbl-pearson-correlations are descriptive summaries
of the same unique off-diagonal pair representation used for the Mantel
analysis. They are included to help interpret the approximate linear agreement
between distance scales, but they are not used as independent-observation
inference.

```{r}
#| label: tbl-pearson-correlations
#| tbl-cap: |
#|   Descriptive Pearson correlations between each distance metric and
#|   ML-tree cophenetic distance, overall and by subtype. Intervals are
#|   strain-level bootstrap intervals and should be interpreted descriptively.

here::here("results", "Tables", "descriptive-pearson-distance-correlation-table.rds") |>
	readr::read_rds() |>
	fit_flextable_to_page()
```

# References

::: {#refs}
:::
